\section{Introduction} \label{sec:Introduction}

A system could be summarised roughly as a solution to one or more problems. One
of the first steps in order to build this kind of solution is to try to
understand the problem -- that is, try to map the requirements of the software.
Vaasen et al. (2009 \cite[cited][p.~8]{Boczko:2012:IAI:2331376}) suggests that
an accounting information system's main purpose is to provide information to
internal and external stakeholders. Although this refers to accounting systems
for businesses, it could be argued that this same definition can be employed to
define personal finance systems -- except that, in this case, the main
stakeholder would be the individual using the system (that is, the user). In
fact, one of the most widely known accounting systems available in the market,
Quicken\texttrademark, was conceived around the idea that there should be more
efficient and less tedious ways to organise one's personal financial
information than doing it manually (\cite{quicken2017about}). This project has
been developed based on similar ideas.

It seems fair to infer that nowadays most of a user's financial transactions
happen in ways that can be listed electronically (usually via their bank or
credit card statements) -- a study by Payments UK
(\citeyear{paymentsUK2017summary}), for example, indicates that there has been
a rise in debit card payments over the past few years, and that the volumes of
this type of transaction is likely to be higher than that of cash payments by
the year 2021. Therefore, an assumption has been made that the users will
require means of uploading a list of their financial transactions into the
system.

The system created for this project intends to do just this. Its main feature,
however, will be to allow the user to categorise expenditure based on patterns
in the entries' descriptions. Aside from this, there will also be a feature to
allow the user to view summaries of the income and expenditure over a period of
time, and another one to generate budget forecasts for future periods based on
the financial information already entered.

This report documents the work of the project. Each chapter delineates a
specific aspect of the development life cycle, which is in line with the
development process listed in Chapter \ref{sec:DevelopmentMethod}: Chapter
\ref{sec:Requirements} outlines the identified requirements which were used as
motivation for the system to be developed; the contents of Chapter
\ref{sec:AnalysisAndDesign} shows further analysis and concomitant design of
the system and the solutions it brings; Chapter \ref{sec:Implementation}
outlines select aspects of the implementation stage which serve to emphasise
the techniques used to implement the designed logic, or highlights areas where
it was felt it was necessary to implement something slightly different that
what was designed.

The system developed for this project has been modelled after the principle of
\emph{double entry bookkeeping}, from the accounting domain, which states that
``money is never created or destroyed -- it merely moves from one account to
another'' (\cite[][Section 6.2]{fowler1997analysis}). More specifically, double
entry is the principle which ensures that every transaction always affects two
accounts, one being credited (Out) and one debited (In). An account, for the
scope of this project, refers either to a category created by the user, or to
the user's \emph{cash book} -- the contents of their bank account plus any
manual entry which they make. In bookkeeping, each account can be classified as
\emph{asset, liability, income} or \emph{expenditure}. Whether the account
increases or decreases will depend on which of these categories it falls under:
\emph{debits} will increase \emph{assets} and \emph{expense} accounts, and
\emph{credits} will increase \emph{liability, capital} or \emph{income}
accounts (\cite[][pp.~18-19]{wood2004book}).
